Optimizing and Incrementalizing Collection Queries by AST Transformations

### 2017-07-25 pldi14/sec-performance.tex
Fix broken ref in performance section, more on self-maintainability
We had something for it to point to already!
Sec.~\ref -> \cref also in pldi14/

\pg{broken reference}
We have seen in \cref{sec:differentiation-fold-example} that $\Derivative$
needlessly recomputes $\Merge\Xs\Ys$. However, the result is a
base input to $\FOLD'$.
%
In next section, we'll replace $\FOLD'$ by a self-maintainable
derivative (based again on $\FOLDBAG$) and will avoid this
recomputation.
\end{examples}

@@

\begin{oldSec} % ID=Appel97
\pg{Do not remove without universal agreement. I don't think the paper can avoid discussing this aspect. }
One important issue is left. In a call-by-value
implementation of lambda calculus, running the program
\[
\Derive{\App{s}{t}} = \App{\App{\Derive{s}}{t}}{\Derive{t}}
\]
computes $t$
again, even though it was computed in the base program, thus
leading to wasteful repeated computation.
However, we claim this is a simpler problem to solve.
Three possibilities arise:
\begin{enumerate}
\item $t$ is very cheap to compute (for instance, it is a
  literal), so the problem does not occur.
\item $t$ is passed to a function which does not use it, hence
  we can avoid computing it using separate optimization steps, by
  executing the derivative with a lazy semantics (as done in our
  benchmarks) or (we expect) by using known techniques for
  interprocedural dead-code elimination~\citep{Appel97}.
\item Otherwise, since the term $t$
  was already computed while running the base
  program, we can save its value for use in the derivative.
  Approaches to
  implement this include memoization~\pg{cite} and extensions of
  static caching (as discussed in Sec.~\ref{sec:rw}).
  We leave investigating the different
  approaches to future work.
\end{enumerate}
\end{oldSec}

@@

Hence, we can argue informally that if a program only uses
self-maintainable primitives, its derivative will not recompute
base intermediate results (except if they are computed for
unrelated reasons). In our case study (Sec.~\ref{sec:eval}),
such derivatives are already
extremely efficient.

>>>>

Similarly to what we have seen in \cref{sec:derive-example-merge} that $\Derivative$
needlessly recomputes $\Merge\Xs\Ys$ without optimizations. However, the result is a
base input to $\FOLD'$.
%
In next section, we'll replace $\FOLD'$ by a self-maintainable
derivative (based again on $\FOLDBAG$) and will avoid this
recomputation.
\end{examples}

@@

\begin{oldSec} % ID=Appel97
\pg{Do not remove without universal agreement. I don't think the paper can avoid discussing this aspect. }
One important issue is left. In a call-by-value
implementation of lambda calculus, running the program
\[
\Derive{\App{s}{t}} = \App{\App{\Derive{s}}{t}}{\Derive{t}}
\]
computes $t$
again, even though it was computed in the base program, thus
leading to wasteful repeated computation.
However, we claim this is a simpler problem to solve.
Three possibilities arise:
\begin{enumerate}
\item $t$ is very cheap to compute (for instance, it is a
  literal), so the problem does not occur.
\item $t$ is passed to a function which does not use it, hence
  we can avoid computing it using separate optimization steps, by
  executing the derivative with a lazy semantics (as done in our
  benchmarks) or (we expect) by using known techniques for
  interprocedural dead-code elimination~\citep{Appel97}.
\item Otherwise, since the term $t$
  was already computed while running the base
  program, we can save its value for use in the derivative.
  Approaches to
  implement this include memoization~\pg{cite} and extensions of
  static caching (as discussed in \cref{sec:rw}).
  We leave investigating the different
  approaches to future work.
\end{enumerate}
\end{oldSec}

@@

Hence, we can argue informally that if a program only uses
self-maintainable primitives, its derivative will not recompute
base intermediate results (except if they are computed for
unrelated reasons). In our case study (\cref{sec:eval}),
such derivatives are already
extremely efficient.
